---
title: "mylab1"
author: "Nathaniel Burola"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rlang)
library(fansi)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

#Read in NOAA_fisheries.csv data 
```{r}
us_landings <- read_csv(here("data", "noaa_fisheries.csv"))
```

#Now make a tidy version of the data frame:
```{r}
landings_tidy <- us_landings %>% 
  janitor::clean_names() %>% #clean_names() converts column names into lowercase 
  mutate(state = str_to_lower(state), #Mutuate will create a new column called state which  is a lowercase version of the state 
         afs_name = str_to_lower(afs_name)) %>% #Straight to lower command will rewrite words
  mutate(dollars_num = parse_number(dollars_usd)) #Parse will drop any non-numeric characters before or after the first number 
```

#Let us get some information for salmon 
```{r}
salmon_landings <- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% # Detect matches for crab
  separate(afs_clean, into = c("group","species"), sep = ",")
  
```

Find some grouped summary data: 

Find annual total US landings and dollar value (summing across all states) for each TYPE of salmon using 'group_by()' + 'summarize()' 

```{r}
salmon_summary <- salmon_landings %>%  
  group_by(year, species) %>%  
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )
```

Make A Graph 
```{r}
salmon_landings_graph <- ggplot(salmon_summary, aes(x = year, y = tot_landings, group = species)) +
  geom_line(aes(color = species)) +
  theme_minimal() +
  labs(x = "year", y = "US commercial salmon landings (pounds)")

salmon_landings_graph
```

Export Your Salmon Value Graph With GGSave 
! Just a reminder that GGsave only applies to the most recent run version of the graph that  was run previously 

```{r}
ggsave(plot = salmon_landings_graph, 
        here("figures", "us_salmon_NB.jpg"), 
        height = 7, 
        width = 7)

#xquartz 
```

Make A Nice Kable Table 
```{r}
salmon_first5 <- salmon_summary %>%  
  head(5)

```

Use Kable : 
```{r}
kable(salmon_first5) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE)
```

